---
title: "Flint Water Data Analysis"
author: "Ilse Paniagua"
date: "10/17/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(magrittr)
library(GGally)
library(prettyR)
```
 
# Establishment

```{r}
load("~/Dropbox/Grad School/UM/GIS/ICPSR_36955/DS0001/36955-0001-Data.rda")

eb <- da36955.0001
```

Keep only variables of interest.

```{r}
eb %<>% select(RESULT_LEAD, NAME, CITY, ZIPCODE)
```

Factor to numeric.
```{r}
lbls <- sort(levels(eb$RESULT_LEAD))
lbls <- (sub("^\\([0-9]+\\) +(.+$)", "\\1", lbls))
eb$RESULT_LEAD <- as.numeric(sub("^\\(0*([0-9]+)\\).+$", "\\1", eb$RESULT_LEAD))
eb$RESULT_LEAD <- add.value.labels(eb$RESULT_LEAD, lbls)
```

Replacing ND with zero.

```{r}
eb[is.na(eb)] <- 0
```


```{r}
#The result is non-detected for lead if the amount of lead in the water was less than 1 ppb. m

#Creating categorical variable
eb$lead_cat <- NA

#1-4
eb$lead_cat[0<= eb$RESULT_LEAD & eb$RESULT_LEAD <=4 ] <- 1

#5 to 14
eb$lead_cat[5<= eb$RESULT_LEAD & eb$RESULT_LEAD <=14 ] <- 2

#15 to 49
eb$lead_cat[15<= eb$RESULT_LEAD & eb$RESULT_LEAD <=49 ] <- 3

#50 to 149
eb$lead_cat[50<= eb$RESULT_LEAD & eb$RESULT_LEAD <=149 ] <- 4

#150 and above
eb$lead_cat[eb$RESULT_LEAD >=150 ] <- 5

```


# Sentinel

```{r}
load("~/Dropbox/Grad School/UM/GIS/ICPSR_36955/DS0002/36955-0002-Data.rda")

sentinel <- da36955.0002
```

Keep only variables of interest.

```{r}
sentinel %<>% select(RESULT_LEAD, NAME = ADDRESS, CITY, ZIPCODE= ZIP_CODE)
```


Replacing ND with zero.

```{r}
sentinel[is.na(sentinel)] <- 0
```

Creating categorical variable.
```{r Sentinel Lead}
#The result is non-detected for lead if the amount of lead in the water was less than 1 ppb. m

#1-4
sentinel$lead_cat[0<= sentinel$RESULT_LEAD & sentinel$RESULT_LEAD <=4 ] <- 1

#5 to 14
sentinel$lead_cat[5<= sentinel$RESULT_LEAD & sentinel$RESULT_LEAD <=14 ] <- 2

#15 to 49
sentinel$lead_cat[15<= sentinel$RESULT_LEAD & sentinel$RESULT_LEAD <=49 ] <- 3

#50 to 149
sentinel$lead_cat[50<= sentinel$RESULT_LEAD & sentinel$RESULT_LEAD <=149 ] <- 4

#150 and above
sentinel$lead_cat[sentinel$RESULT_LEAD >=150 ] <- 5
```


# School

```{r}
# School
load("~/Dropbox/Grad School/UM/GIS/ICPSR_36955/DS0003/36955-0003-Data.rda")

schools <- da36955.0003
```

Factor to numeric.
```{r}
lbls <- sort(levels(schools$RESULT_LEAD))
lbls <- (sub("^\\([0-9]+\\) +(.+$)", "\\1", lbls))
schools$RESULT_LEAD <- as.numeric(sub("^\\(0*([0-9]+)\\).+$", "\\1", schools$RESULT_LEAD))
schools$RESULT_LEAD <- add.value.labels(schools$RESULT_LEAD, lbls)
```

```{r}
schools %>%
  group_by(SITEDESCRIPTION) %>%
  summarise(count = n(), mean((RESULT_LEAD))) %>%
  arrange(desc(count))
```

Keeping only post-filter installments.
Treating observations within a school as independent (keeping all observations).

```{r}
#Cleaning site description variable
schools$SITEDESCRIPTION <- str_replace_all(schools$SITEDESCRIPTION,"                                ","")
schools$SITEDESCRIPTION <-str_replace_all(schools$SITEDESCRIPTION,"                               ","")
schools$SITEDESCRIPTION <-str_replace_all(schools$SITEDESCRIPTION,"  ","")

#Filtering

schools %<>% filter(TESTINGTYPE !="(1) Pre-Fixture Replacement") %>% select(RESULT_LEAD, NAME = SCH_NAME, CITY = SCH_ZIP, ZIPCODE = SCH_ZIP)
```

Replacing NA with zero.

```{r}
schools[is.na(schools)] <- 0
```


```{r School Lead}
#The result is non-detected for lead if the amount of lead in the water was less than 1 ppb.

#Creating categorical variable

#1-4
schools$lead_cat[0<= schools$RESULT_LEAD & schools$RESULT_LEAD <=4 ] <- 1

#5 to 14
schools$lead_cat[5<= schools$RESULT_LEAD & schools$RESULT_LEAD <=14 ] <- 2

#15 to 49
schools$lead_cat[15<= schools$RESULT_LEAD & schools$RESULT_LEAD <=49 ] <- 3

#50 to 149
schools$lead_cat[50<= schools$RESULT_LEAD & schools$RESULT_LEAD <=149 ] <- 4

#150 and above
schools$lead_cat[schools$RESULT_LEAD >=150 ] <- 5
```


#Tables
Number of schools, establishments, sentinel
Mean level of lead for each group
Proportion of observations above 15 ppb


```{r}
#create identifer for each group
 schools %<>% mutate(type = "School", CITY="Flint")
 sentinel %<>% mutate(type = "Sentinel")
  eb %<>% mutate(type = "Establishment")
```

```{r}
#Append records

test <- full_join(schools, sentinel)

test <- full_join(test, eb)
```

```{r}
test %>%
  mutate(prop= ifelse(RESULT_LEAD >=15,1,0)) %>%
  group_by(type) %>%
  summarise(Records= n(), "Mean Lead"= mean(RESULT_LEAD), "% High Lead" = (mean(prop) * 100), "Max Lead" = max(RESULT_LEAD))
```

```{r}
#Histogram
na.omit(test) %>%
  ggplot(aes(x=RESULT_LEAD)) +
  geom_histogram(aes(fill=type),binwidth=2.5) +
  facet_grid(~type, scales="free") +
  lims(x= c(0,100), y= c(0,200)) +
  theme_light() +
  scale_fill_brewer(palette="Dark2") +
  labs(title="Histogram of Lead Results",
       x="Lead (ppb)",
       y="Count",
       fill="Type") +
  theme(strip.background =element_rect(fill="darkgrey"))
```


Write final CSV file.

```{r}
write.csv(test, file="dataqualitypoints_13nov.csv")
```

# Create correlation matrix

References:
https://www.r-graph-gallery.com/199-correlation-matrix-with-ggally.html

Reading table generated from ArcGIS from a spatial join between all the data quality points and their census tracts.

```{r}
acs <- read.csv("ACS_Flint_SpatialJoin_TableToExcel.csv")
```

```{r}
#Overall
info <- acs %>%
  mutate(highpb = ifelse(lead_cat >=3,1,0)) %>%
  select(Est..Black.or.Af..Amer..alone, Est..White.alone, Est..Total.1, Est.....10K, BelowPoverty, highpb, lead_cat, Est..Med..Year)

names(info) <- c("BlackHH", "WhiteHH","TotalPop", "Housing10k","BelowPvty", "HighLead","lead_cat", "HHAge")

info %<>%
  group_by(BlackHH, WhiteHH, TotalPop, Housing10k, BelowPvty, HHAge) %>%
  mutate(highpb = ifelse(lead_cat >=3,1,0)) %>%
  summarise(HighLead = mean(highpb, na.omit=TRUE)) %>%
  na.omit()
  
```


```{r}
info %>%
  ggcorr(., method= c("everything", "pearson"),label=TRUE,label_round=2,layout.exp=2.5) +
  theme(plot.title= element_text(size=15, hjust=0.75))

```

